\chapter{Introduction}%
\label{chapter:introduction}

\begin{introduction}
This chapter summarizes the challenges and significance of integrating Wi-Fi-only devices into the 5G core network, highlighting the limitations of current standards and the need for innovative solutions to enable seamless connectivity and authentication.
\end{introduction}

\section{Background and Context}

% Introduce 5G as a significant advancement in wireless technology, highlighting its core features such as higher speeds, lower latency, and support for diverse use cases (e.g., enhanced mobile broadband, massive machine-type communications, ultra-reliable low-latency communications).

In recent years, \ac{5G} wireless technology has revolutionized telecommunications. It offers higher bandwidth, faster speeds, and lower delays, supporting use cases like \ac{eMBB}, \ac{mMTC}, and \ac{uRLLC}.

Additionally, it is transforming private networks, which have traditionally relied on legacy wired or wireless (WLAN) Ethernet. Features like tighter security, higher reliability, and \ac{TSN} are crucial to meeting Industry 4.0 requirements for wireless connectivity. However, how can the industry bridge the gap between existing \ac{N5GC} devices, which current networks rely on, and new \acp{5GC}?

% Briefly describe the Service Based Architecture of 5G, emphasizing the incorporation of Network Functions Virtualization (NFV) and Software Defined Networking (SDN).
% Mention the various access nodes supported by the 5G system, including native 5G-NR and LTE accesses, and non-3GPP interworking functions.

\ac{5G} is not only a revolution in radio network infrastructure but also in the core network. Now based on Service Based Architecture, using \ac{NFV} and \ac{SDN}, it allows it to minimize cost and maximize utilization and elasticity of the infrastructure by seperating the \ac{UP} functions from the \ac{CP} function. This architecture supports various access nodes such as native \ac{NR}, LTE accesses, and non-\ac{3GPP} interworking functions that facilitate connectivity from untrusted WLANs.~\cite{23.501-p41}

% Discuss the importance of integrating Wi-Fi networks with 5G to enhance connectivity options for users and devices.
% Highlight the challenges posed by current standards that do not support Wi-Fi-only devices without USIM in connecting to the 5G Core.

As wireless networks evolve, the convergence of 5G with existing Wi-Fi infrastructures becomes increasingly critical. However, current standards established by the \ac{3GPP} do not adequately address the integration of Wi-Fi-only devices that lack \ac{USIM} capabilities into the \ac{5GC} network.~\cite{wba-04-2021-p59} This limitation is particularly evident in enterprise environments where many devices operate solely on Wi-Fi. To fully realize the potential of  technology, it is essential to develop solutions that enable integration of Wi-Fi-only devices into the \ac{5G} ecosystem. This includes addressing challenges related to authentication mechanisms, device identity, and overall interoperability between different network types.

\section{Problem Statement}

% - Highlight that the current 3GPP standard does not define an architecture to support Wi-Fi-only devices without USIM connecting to the 5G Core.
% - Emphasize the importance of this issue, especially in enterprise deployments where many devices lack USIM capabilities.
% - Mention the Wireless Broadband Alliance (WBA) recommendation for 3GPP to define procedures for supporting Wi-Fi-only UE with non-IMSI based identity and EAP-TLS/EAP-TTLS based authentication.

The current \ac{3GPP} standard lacks an architecture for integrating Wi-Fi-only devices without \ac{USIM} into the \acl{5GC}, creating a significant gap in connectivity. This limitation is problematic in enterprise environments, where many devices operate only on Wi-Fi and do not possess \ac{USIM} capabilities. The \ac{WBA} has identified this issue, recommending that \ac{3GPP} develop procedures to support Wi-Fi-only \ac{UE} using non-\acs{IMSI} based identity and authentication methods such as \ac{EAP-TLS} or \ac{EAP-TTLS}. Addressing this challenge is crucial for enabling integration of diverse device types into the \ac{5G} ecosystem.

\section{Research Objectives}

% - State your primary goal: to explore and develop solutions for integrating Wi-Fi-only devices without USIM into the 5G Core infrastructure.
% - List specific objectives, such as:
%   - Investigating authentication mechanisms compatible with both 5G and Wi-Fi networks.
%   - Developing a method for device identification that works across 5G and non-3GPP networks.
%   - Proposing extensions or alternatives to existing protocols to support seamless integration.

The main goal of this research is to explore and develop solutions for integrating Wi-Fi-only devices without \ac{USIM} into the \ac{5GC} infrastructure. To achieve this, we have identified the following specific objectives:

\begin{enumerate}
    \item{
        Investigate authentication mechanisms compatible with both \ac{5G} and Wi-Fi networks:
        \begin{itemize}
            \item Analyze existing authentication methods such as \ac{EAP-TLS} and \ac{EAP-TTLS} for their applicability in a converged \ac{5G}-Wi-Fi environment.
            \item Explore potential modifications or extensions to these methods to ensure seamless authentication across different network types.
        \end{itemize}
    }
    \item{
        Develop a method for managing device identity that works across \ac{5G} and non-\ac{3GPP} networks:
        \begin{itemize}
            \item Investigate the possibility of designing an extended \ac{NAI} or alternative identifier that can accommodate Wi-Fi-only devices while maintaining compatibility with the \ac{5G} infrastructure.
            \item Investigate the possibility of generating a pseudo-\ac{SUCI} and pseudo-\ac{SUPI} for \ac{N5GC} devices that follows the \ac{NAI} format and serves a similar function in the authentication flow.
        \end{itemize}
    }
    \item{
        Propose extensions or alternatives to existing protocols:
        \begin{itemize}
            \item Investigate the possibility for mapping existing \ac{N5GC} device identifiers (e.g., \ac{MAC} addresses) to a format compatible with the \ac{5G} authentication framework.
            \item Explore the potential for enhancing or creating new \ac{EAP} methods specifically designed for \ac{N5GC} devices in a \ac{5G} context.
        \end{itemize}
    }
\end{enumerate}

\section{Thesis Structure}

This document explores the challenge of integrating Wi-Fi-only devices into the \ac{5GC}. We begin by examining the current landscape of \ac{5G} and Wi-Fi integration, focusing on authentication mechanisms and their limitations. Building on this foundation, we propose a framework to address these challenges, detailing our approach to authentication and device identification. We then put our solution to the test, presenting experimental results and comparing them with existing methods. Finally, we reflect on our contributions, acknowledge the boundaries of our work, and suggest possible avenues for future research. Through this journey, we aim to contribute meaningfully to the ongoing convergence of \ac{5G} and Wi-Fi technologies.